### 2020-08-23 thesis/thesis.tex
[thesis] Abstract

\polishtitle{Efekty koalgebraiczne oraz ich kohandlery \fmlinebreak{} w językach programowania}
\englishtitle{Coalgebraic effects and their cohandlers \fmlinebreak{} in programming languages}
\polishabstract{
    \ldots
}
\englishabstract{
    \ldots
}
\author{Mateusz Urbańczyk}
\advisor{dr Maciej Piróg}
\date{1 września 2020}                     % Data zlozenia pracy
% Dane do oswiadczenia o autorskim wykonaniu
\transcriptnum{291480}                     % Numer indeksu
\advisorgen{dr. Macieja Piróga} % Nazwisko promotora w dopelniaczu
%%%%%

@@

In this thesis we discuss issues arising from combining algebraic effects
and handlers wiht resource management, and present a solution through coalgebraic
means along with presentation of experimental programming language Freak, which
is a language with (co)algebraic effects and (co)handlers, where implementation
is based on Continuation Passing Style translation.

>>>>

\polishtitle{Efekty koalgebraiczne oraz ich kohandlery \fmlinebreak{} w językach programowania}
\englishtitle{Coalgebraic effects and their cohandlers \fmlinebreak{} in programming languages}
\polishabstract{
    Efekty algebraiczne są dobrym rozwiązaniem na modelowanie efektów
    obliczeniowych, w sposób który dobrze się składa. Niemniej, powstają przy
    nich problemy z idiomatycznym zarządzaniem zasobami. Z matematycznej
    strony, tworzą one wolne modele nad teoriami algebraicznymi, a ich handlery
    homomorfizmy zachowujące strukturę modelu. Dualnym pojęciem są komodele,
    które okazały się odpowiednim narzędziem do wyrażania w kodzie przejść z
    przekazywaniem stanu. Prezentujemy rachunek z efektami algebraicznymi i
    koalgebraicznymi, aby zaadresować problemy z zarządzaniem zewnętrznym stanem.
}
\englishabstract{
    Algebraic effects are a great way to model computational effects in a
    composable way. Nonetheless, they give rise to issues with idiomatic
    resource management. From mathematical side, they form
    free models over algebraic theories, with handlers being homomorphisms
    preserving the model structure. Dualizing this notion gives us comodels,
    which turned out to be suitable for expressing state-passing transitions
    in the code. We present a calculus with algebraic and coalgebraic effects,
    to address the issues with external state management and provide abstraction
    over resources.
}
\author{Mateusz Urbańczyk}
\advisor{dr Maciej Piróg}
\date{1 września 2020}                     % Data zlozenia pracy
% Dane do oswiadczenia o autorskim wykonaniu
\transcriptnum{291480}                     % Numer indeksu
\advisorgen{dr. Macieja Piróga} % Nazwisko promotora w dopelniaczu
%%%%%

@@

In this thesis we discuss issues arising from combining algebraic effects
and handlers with resource management, and present a solution through coalgebraic
means along with presentation of experimental programming language Freak, which
is a language with (co)algebraic effects and (co)handlers, where implementation
is based on Continuation Passing Style translation.
